Simulation connectance richness

### 2023-05-19 _targets.Rmd
Sim with higher persistence hopefully: low Z and K gradient

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse", "magrittr", "cowplot", "here", "arrow", "piecewiseSEM", "semEff", "easystats", "glmmTMB"))
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
source(here::here("R", "model_helpers.R"))
```

@@

@@

```{r, message = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(biglm)
```

>>>>

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse", "magrittr", "cowplot", "here", "arrow", "piecewiseSEM", "semEff", "easystats", "glmmTMB"))
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
source(here::here("R", "stability_decomposition.R"))
source(here::here("R", "model_helpers.R"))
```

@@

```{targets raw_sim}
list(
  tar_target(var_param_sim, c("rep", "productivity", "S", "ct", "rho", "env_stoch", "Z")),
  tar_target(sim_files, list.files(here::here("res", "simCSZenrich2/simCSZenrich2"))),
  tar_target(sim_ts,
    map_dfr(sim_files,
      function (x) {
        open_dataset(here::here("res", "simCSZenrich2/simCSZenrich2", x),
          format = "arrow") %>%
        collect() %>%
      filter(!is.na(richness)) %>%
      select(c(var_param_sim, "species", "stoch")) %>%
      # Transform back interaction strength into matrices
      mutate(across(where(is.list), as.list)) %>%
      mutate(
        stoch_d = map2(species, stoch, function(bm, d_stoch) {
          .1 * bm * exp(d_stoch)
          }),
        species = map(species, ~matrix(.x, nrow = 100)),
        stoch = map(stoch, ~matrix(.x, nrow = 100)),
        stoch_d = map(stoch_d, ~matrix(.x, nrow = 100)),
        cpe = map_dbl(species, compensatory_effect),
        cpe_env = map_dbl(stoch_d, compensatory_effect),
        cpe_int = cpe / cpe_env,
        stab_com = map_dbl(species, community_stability),
        pop_stab = map_dbl(species, population_stability),
        async = map_dbl(species, asynchrony),
        sae = map(species, statistical_averaging_effect)
      )
      })
  ),
  tar_target(sim_stab, sim_ts %>% select(-species, -stoch_d, -stoch)),
  tar_target(sim_prep,
map_dfr(sim_files,
      function (x) {
        open_dataset(here::here("res", "simCSZenrich2/simCSZenrich2", x),
          format = "arrow") %>%
        collect() %>%
      filter(!is.na(richness)) %>%
      select(-species, -stoch, -A) %>%
      # Transform back interaction strength into matrices
      mutate(across(where(is.list), as.list)) %>%
      mutate(across(c(int_strength, max_int),
          ~map2(.x, S,
            function(y, rich) {
              matrix(y, nrow = rich)
            }
          )
      )
        ) %>%
      # Get matrix of alive species
      mutate(max_int_alive  = map2(max_int, bm_sp,
          function(mat, bm) {
            mask <- bm > 10^-6
            mat[mask, mask, drop = FALSE]
          }
        )
      )
      }

    )),
  tar_target(sim,
    sim_prep %>%
      mutate(
        persistence = richness / S,
        async = 1 / sync,
        stab_pop = 1 / avg_cv_sp,
        ct_alive = map_dbl(max_int_alive,
          ~ifelse(sum(.x) == 0, 0,
            sum(.x > 0) / ((ncol(.x))^2))),
        bm_sp_alive = map(bm_sp, ~.x[.x > 10^-6]),
        tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-6]),
        max_tlvl_alive = map_dbl(tlvl_alive, max),
        w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive,
          ~sum(.x * (.y / sum(.y)))
          ),
        max_max_int_alive = map_dbl(max_int_alive,
          ~ifelse(length(.x[.x > 0]) == 0, 0, max(.x[.x > 0]))
          ),
        avg_max_int_alive = map_dbl(max_int_alive,
          ~ifelse(length(.x[.x > 0]) == 0, 0, mean(.x[.x > 0]))
          ),
        sd_max_int_alive = map_dbl(max_int_alive,
          ~ifelse(length(.x[.x > 0]) <= 1, 0, sd(.x[.x > 0]))
          ),
        cv_max_int_alive = sd_max_int_alive / avg_max_int_alive,
        inv_sd_max_int_alive = 1 / sd_max_int_alive,
        gini_max_int_alive = map_dbl(max_int_alive,
          ~ifelse(length(.x[.x > 0]) == 0, 0, gini(.x[.x > 0]))),
        avg_omnivory = map_dbl(omnivory, mean),
        disconnected_prod = map_lgl(max_int_alive,
        ~any(colSums(.x) == 0 & rowSums(.x) == 0))
      ) %>%
    select(!where(is.list))
  )
)
```

@@

```{r, message = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
```

```{r}
tar_read(sim_ts)
```

### 2023-05-19 report/cs_analysis.Rmd
Sim with higher persistence hopefully: low Z and K gradient

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
---

@@

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(ggcorrplot)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(targets)

@@

```{r, eval = FALSE}
colnames(sim)
unique(simZ$env_stoch)
summary(simZ$total_biomass)
summary(simZ$persistence)
summary(simZ$ct_alive)
summary(simZ$gini_max_int_alive)
```

@@

- Response diversity is key in driving stability in ecological communities
- Response diversity control asynchrony with richness
- Mainly done in plants or in same food-web modules
- Food-webs structure can also modulate synchrony and population stability by
  the strength of trophic interactions

@@

@@

>>>>

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
always_allow_html: true
bibliography: references.bib
---

@@

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(ggcorrplot)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)

@@

@@

Understanding the drivers of temporal stability is a key question in Ecology.
BEF theory shows that species richness is a key driver to understand the
temporal stability of community biomass.

Species richness was demonstrated to increase community stability by driving higher
asynchrony in community in various assemblages and ecosystems. In particular,
high response diversity, which drives independant species fluctuations, seems to a key
factor driving the positive effect of species richness of stability
[@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022]. However,
those studies were often limited to monotrophic assemblages. As natural
communities are often composed of multiple trophic level, our understanding of
community stability needs to incorporate food-web structure.

Food-web structure has been shown to have tremendous effect on community
stability. The degree of generalism of predator was shown to modulate
stability-richness relationships [@thebault_trophic_2005]. The strength and distribution of
interaction strength was also shown to be strong driver of community stability
[@mccann_weak_1998]. Omnivory was also shown to drive higher stability
[@mccann_reevaluating_1997]. However, it is quite unkown how those food-web
structure elements would affect stability-species richness relationships.

The effect of environmental stressors can also strongly affect community
variability. Higher environmental stress and variability may dampens the effect
of richness on stability [@loreau_biodiversity_2001]. Lastly, productivity of a
system, as the quantity of biomass input in the ecosystem, can have tremendous
effect on both community structure and stabilty [@lindeman_trophic-dynamic_1942;
@binzer_dynamics_2012; @binzer_interactive_2016]. Productivity can increase
species richness in the food-web by increasing the level of energy input
[@lindeman_trophic-dynamic_1942] but also destabilises community biomass by
increasing population variability, a behavior coined as the paradox of
enrichment [@rosenzweig_paradox_1971]. A better understanding of the effect of
species richness on community stability should then include environmental
gradient and stressors.

Community stability can itself be partitioned in several components, i.e.
population stability and population asynchrony [@thibaut_understanding_2013;
@zhao_biodiversity_2022]. Population stability can be itself partioned in total
biomass and population variability effect. Population asynchrony can further be
partioned in statistical averaging effect and compensatory dynamics effects
[@zhao_biodiversity_2022]. Partitioning stability allows to better understand
how food-web structure and external factors affect stability. For exemple,
environmental stochasticity is expected to primarily affect population
stability, while response diversity should affect primarily compensatory
dynamics.

```{r}
concept_cap <- "Food-web structure, response diversity and enviromental gradient
can modulate stability-richness relationships. (a & b) Food-web structure may
modulate stability-richness relationships additively (a) or interactively (b).
(c) Decrease in response diversity and increase in environmental variability
should dampen stability-richness relationships. While the loss of response
diversity should dampen the effect of species richness on asynchrony, the
increase in environmental variability should decrease primarily population
stability. (d) Increase in productivity should increase species richness
(trophic energy hypothesis) but also dampens the relationship between stability
and species richness (paradox of enrichment). (e) Summary of the framework, how
food-web structure and external perturbations modulate the effects of species
richness on community stability. More precisely, how they affect the different
components of stability: total biomass, population variability, statistical
averaging effects and compensatory dynamics."
```

```{r, fig.cap=concept_cap}
knitr::include_graphics(here("report", "figures", "conceptual_figures.pdf"))
```

```{r}
tribble(
  ~Type, ~`Food-web structure`, ~`Effect on stability-richness`,
  "Food-web structure", "Connectance", "Strengthen (Miele et al., 2016) / No effect (Thébault & Loreau, 2005)",
  "Food-web structure", "Omnivory", "Strengthen (Mccann et al., 1997)",
  "Food-web structure", "Interaction strength", "Destabilising (Mccann et al., 1998) / Stabilising in
  productive (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Response diversity", "", "Dampens",
  "Productivity", "", "Dampens (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Environmental variability", "", "Dampens",
)
```

```{r, eval = FALSE}
tar_load(sim_prep)
# test for disconnected species
ti <- sim_prep %>%
  filter(max_tlvl > 1) %>%
  mutate(
    disconnected_prod = map_lgl(max_int_alive,
      # which species has no predator (colsum) is not a predator
      ~any(colSums(.x) == 0 & rowSums(.x) == 0)
    )
  )

ti %>%
  ggplot(aes(x = as.factor(disconnected_prod), y = stab_com)) +
  geom_boxplot()
```

```{r}
tar_load(sim_stab)
sim_stab %<>%
  mutate(
    sae_even = map_dbl(sae, ~.x["even"]),
    evenness_sae = map_dbl(sae, ~.x["eveness"]),
    sae_total = map_dbl(sae, ~.x["total"])
  )
tar_load(sim)
sim_ok <- sim %>%
  select(-stab_com, -async) %>%
  left_join(sim_stab) %>%
  mutate(inv_sum_pop_var = pop_stab / total_biomass)
```

```{r}
sim_ok %>%
  filter(ct_alive != 0, rho == 0, env_stoch == .1)
sim_sem <- sim_ok %>%
  filter(ct_alive != 0) %>%
  mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}")) %>%
  na.omit()
```

```{r}
sem_list_basic <- list(
  w_avg_tlvl_alive = lm(w_avg_tlvl_alive ~ richness + ct_alive, sim_sem),
#  avg_omnivory = lm(avg_omnivory ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem),
  sd_max_int_alive = lm(sd_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive + Z, sim_sem),
  avg_max_int_alive = lm(avg_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive + Z, sim_sem),
  log_inv_sum_pop_var = lm(log_inv_sum_pop_var ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + env_stoch + rho, sim_sem),
  log_total_biomass = lm(log_total_biomass ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + env_stoch + rho, sim_sem),
  log_sae_even = lm(log_sae_even ~ richness, sim_sem),
  log_evenness_sae = lm(log_evenness_sae ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_cpe_env = lm(log_cpe_env ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + rho + env_stoch, sim_sem),
  log_cpe_int = lm(log_cpe_int ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_async = lm(log_async ~ log_sae_even + log_evenness_sae + log_cpe_env + log_cpe_int, sim_sem),
  #log_stab_pop = lm(log_pop_stab ~ log_total_biomass + log_inv_sum_pop_var, sim_sem),
  log_stab_com = lm(log_stab_com ~ log_total_biomass + log_inv_sum_pop_var + log_async, sim_sem)
)
sem_basic <- as.psem(sem_list_basic)
ti <- summary(sem_basic)$coefficients[, c(1, 2, 8, 9)]
colnames(ti)[4] <- "Signif"
ti
```
```{r, eval = FALSE}
lm(
  log_stab_com ~ log_total_biomass + log_inv_sum_pop_var + log_async,
  sim_ok %>%
    filter(ct_alive == 0) %>%
    mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}"))%>%
    mutate(across(where(is.double), ~scale(.x)[,1]))
)
```

```{r}
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = pop_stab)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = pop_stab)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
```

```{r}
sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(y = avg_omnivory, x = richness)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(y = avg_omnivory, x = ct_alive)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = avg_omnivory, y = 1/avg_cv_sp)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = avg_omnivory, y = sync)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = Z, y = max_tlvl_alive)) +
  geom_point()
```

```{r}
p_sim_sync_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = sync, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))

p_sim_pop_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = 1/avg_cv_sp, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))
```

```{r}
p_sim_stab_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = stab_com, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))
ggsave(
  filename = here("report", "figures", "sim_stab_richnness_rho_stoch.pdf"),
  p_sim_stab_rich,
  width = 80,
  height = 80 * 4,
  units = "mm",
  scale = 2.4
)
```

```{r}
knitr::include_graphics(here::here("report", "figures", "sim_stab_richnness_rho_stoch.pdf"))
```

# Annexes

@@

- Brose started with 5 basal species and K = 1, so his minimum total K is 5

@@

# Appendix

```{r}
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = max_tlvl_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = max_tlvl_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = avg_max_int_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(productivity))

sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = persistence, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = pop_stab, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = ct_alive, y = persistence, color = as.factor(S))) +
  geom_point()
```

- K = 1 is too low for meaningful food-webs, I should start with K = 5 as in
  Brose et al. (2006)
